---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%",
  dpi = 400,
  fig.align = "center"
)
```

# ididvar

<!-- badges: start -->
<!-- badges: end -->

This package provides tools to easily **identify the identifying variation in a regression**, specifically in **applied economics** analyses. It does not provide new metrics per say but instead facilitates the implementation of slightly modified versions of existing ones.

This package is built as part of a [research project](https://vincentbagilet.github.io/causal_exaggeration/). As such, the [associated paper](https://vincentbagilet.github.io/causal_exaggeration/causal_exaggeration_paper.pdf) provides a detailed scientific description of its content and of its underpinnings. 

## Installation

You can install the development version of `ididvar` from [GitHub](https://github.com/vincentbagilet/ididvar) with:

```{r instal, eval=FALSE}
# install.packages("devtools")
devtools::install_github("vincentbagilet/ididvar")
```

## Usage

Most of the functions of the package take as input the output of a regression and the name of the variable of interest. `ididvar` supports a breadth of estimation packages (`lm`, `plm`, `fixest` for instance).

First, the package provides a straightforward function to compute identifying variation weights. 

```{r compute_weights, message=FALSE}
library(ididvar)

reg_ex_fixest <- ggplot2::txhousing  |>
  fixest::feols(fml = volume ~ sales + listings |  as.factor(date) + city)

idid_weights(reg_ex_fixest, "sales") |>
  head()
```

The package also allows for an effortless exploration of these weights through visualizations.

```{r plot_weights, message=FALSE, fig.asp=1.5, fig.dpi=400}
library(ggplot2)

idid_viz_weights(reg_ex_fixest, "sales", date, city) +
  ggplot2::labs(x = NULL, y = NULL) 
```

In this particular (somehow silly) regression, most of the identifying variation comes from a few cities: Houston, Dallas, Austin and San Antonio.

```{r plot_cumul, message=FALSE, fig.asp=1.3, fig.dpi=400, out.width='60%'}
idid_weights(reg_ex_fixest, "sales") |>
  idid_viz_cumul()
```

It also provide functions to identify observations that actually contribute to identification (in the sense that dropping the other observations does not significantly affect the estimate obtained). 

```{r plot_contrib, message=FALSE, fig.asp=1.5, fig.dpi=400}
idid_viz_contrib(reg_ex_fixest, "sales", city) +
  ggplot2::labs(x = NULL) +
  ggplot2::coord_flip()
```

The package provides a larger set of functions.The [Get started vignette](articles/ididvar.html) introduces them in a concise manner, while also describing a typical workflow for analysis. 

[Online appendices](https://vincentbagilet.github.io/causal_exaggeration/) of the [associated paper](https://vincentbagilet.github.io/causal_exaggeration/causal_exaggeration_paper.pdf) complements this vignette by providing a thorough example of a practical implementation of an analysis using the `ididvar` package. 

